Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-21 paper/sections/proof-reconstruction.tex
[ paper ] working.

The proof-reconstruction approach proposed here consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is the proof
assistant, \Agda. We choose \Agda, but a proof assistant with
support of type theory and similar features like \Agda
(see a description of \Agda in Section
\ref{sec:implementation}) can be used.
Following the diagram in Fig.
\ref{fig:proof-reconstruction-workflow}, the process can begin with
a problem in \CPL that we encode in \TPTP format to be sent as the
input of the \Metis prover.
The algorithms of \Metis in Step~3 intend to find
a solution; if they success, \Metis replies a derivation in \TSTP
format of the proof, else the problem is counter feasible.

@@

\begin{equation}
\label{eq:unshunt-definition}
\unshunt_{0}(φ) =
\begin{cases}
\unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\

\unshunt_{0}(φ₁ ⇒ φ₂) ∧ \unshunt_{0}(φ₁ ⇒ φ₃)
  & φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃),\\

φ &\text{otherwise}.
\end{cases}
\end{equation}

@@

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
\unshunt_{1}(φ, \zero) &= φ\\
\unshunt_{1}(φ, \suc(n)) &=
\begin{cases}
\unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n)
  & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\unshunt_{1}(φ₁ ⇒ φ₂, n)  & φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).\\
∧\ \unshunt_{1}(φ₁ ⇒ φ₃, n)
\end{cases}
\end{split}
\end{align}

@@

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
rank(φ₁) + c_{¬}  & φ ≡ \neg φ₁, \\
\max{(rank(φ₁),rank(φ₂))} + c_{\square}
  & φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

@@

\begin{equation}
\label{eq:unshunt-rank-definition}
\unshunt_{cm}(φ)=
\begin{cases}
\unshunt_{cm}(φ₃) + 2 & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\max(\unshunt_{cm}(φ₂),\unshunt_{cm}(φ₃)) + 1
  & φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃),\\
0 &\text{otherwise.}
\end{cases}
\end{equation}

@@

\emph{Splitting a goal.}
The goal of a proof is processed by \Metis to generated new targets,
\Metis calls these new formulas \emph{subgoals}.
Then, \Metis proves each subgoal and with that it finishes the proof
of the goal.

Nevertheless, \Metis does not provide neither an explanation of this
process in its \TSTP derivations nor in its documentation.
We provide such a evidence that the process of creating subgoals,
and combining them in a final conjunction proves the original goal.

\begin{theorem}[\thmunshunt]
\label{thm:unshunt}
$Γ ⊢ \unshunt(φ)$, then $Γ ⊢ φ$.
\end{theorem}

@@

Now, let us consider the \mysplit function that
yields the conjunction of subgoals from a proposition goal.

\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
\unshunt(\mysplit(φ₁))              & φ ≡ φ₁ ∧ φ₂,\\
∧\ \unshunt(φ₁ ⇒ \mysplit(φ₂))      &\\[2mm]

\unshunt(¬ φ₁ ⇒ \mysplit(φ₂))       & φ ≡ φ₁ ∨ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂))         & φ ≡ φ₁ ⇒ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂))         & φ ≡ φ₁ ⇔ φ₂,\\
∧\ \unshunt(φ₂ ⇒ \mysplit(φ₁))      &\\[2mm]

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))       & φ ≡ ¬ (φ₁ ∧ φ₂),\\

\unshunt(\mysplit(¬ φ₁))            & φ ≡ ¬ (φ₁ ∨ φ₂),\\
∧\ \unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂))  &\\[2mm]

\unshunt(\mysplit(φ₁))              & φ ≡ ¬ (φ₁ ⇒ φ₂),\\
∧\ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂))    &\\[2mm]

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))       & φ ≡ ¬ (φ₁ ⇔ φ₂),\\
∧\ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))    &\\[2mm]

\unshunt(\mysplit(φ₁))              & φ ≡ ¬ ¬ φ₁,\\
⊤                                   & φ ≡ ¬ ⊥,\\
⊥                                   & φ ≡ ¬ ⊤.
\end{cases}
\end{equation}

We reformulate \mysplit function to match with the
new definition after apply the process described above in
Subsection~\ref{ssec:structural-recursion}, since
in the recursion calls of \mysplit in Eq.~\ref{eq:split-definition},
the negations, $(¬ φ_{1,2})$ raise a terminating problem.

\begin{theorem}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\end{theorem}

@@

(φ ≡ φ₁ ∧ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\UnaryInfC{Γ ⊢ φ₁}

\AxiomC{$\mathcal{D}_1$}
\UnaryInfC{Γ ⊢ φ₁}

\AxiomC{$Γ ⊢ \mysplit(φ,n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{$∧$-proj$₂$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{$⇒$-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ φ₂$}

\RightLabel{∧-intro}
\BinaryInfC{$Γ ⊢ φ₁ ∧ φ₂$.}
\end{bprooftree}
\end{equation*}
Where,
\begin{equation*}
(\mathcal{D}_1)\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ,n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁$.}
\end{bprooftree}
\end{equation*}

(φ ≡ φ₁ ⇒ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{}
\RightLabel{assume φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
\AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{weaken φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{⇒-elim}
\BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
\RightLabel{⇒-intro}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$.}
\end{bprooftree}
\end{equation*}

(φ ≡ ¬ (φ₁ ∨ φ₂))
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

\AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂$}
\RightLabel{∧-intro}
\BinaryInfC{$Γ ⊢ ¬ φ₁ ∧ ¬ φ₂$}
\RightLabel{¬∧¬-to-¬∨}
\UnaryInfC{$Γ ⊢ ¬ (φ₁ ∨ φ₂)$.}
\end{bprooftree}
\end{equation*}

Where,

\begin{equation*}
(\mathcal{D})
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$.}
\end{bprooftree}
\end{equation*}
\end{proof}

\begin{theorem} If $s_i$ are the subgoals for the goal φ, and
$\{ Γ ⊢ s_i \}_{i}$, then $Γ ⊢ φ$.
\end{theorem}

@@

\begin{align}
\begin{split}
conjunct(φ, ψ) =
\begin{cases}
ψ &\text{if }φ ≡ ψ\\
ψ &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }ψ ≡ conjunct(φ₁, ψ)\\
ψ &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }ψ ≡ conjunct(φ₂, ψ)\\
φ &\text{otherwise}
\end{cases}
\end{split}
\end{align}

@@

\begin{proof}
When φ ≡ ψ, from $Γ ⊢ conjunct(φ, ψ)$ by normalization, we get $Γ ⊢ ψ$.
By using $subst$ theorem Eq.~\ref{eq:substitution-theorem}, we get the desire conclusion. If the proposition φ is a conjunction, and $ψ ≡ conjunct(φ_{i}, ψ)$
for some $i = 1,\ 2$, from $Γ ⊢ conjunct(φ, ψ)$ by normalization, we get
$Γ ⊢ ψ$, but since $conjunct(φ_{i}, ψ) ≡ ψ$; we apply $subst$ theorem again,
and the theorem follows.
For the latter case, by using the hypothesis, $Γ ⊢ φ$, the proof is complete.

>>>>

The proof-reconstruction approach proposed here consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system are the proof
assistant, \Agda. We choose \Agda, but a proof assistant with
support of type theory and similar features like \Agda
(see a description of \Agda in Section
\ref{sec:implementation}) can be used.
Following the diagram in Fig.
\ref{fig:proof-reconstruction-workflow}, the process can begin with
a problem in \CPL that we encode in \TPTP format to be sent as the
input of the \Metis prover.
The algorithms of \Metis in Step~3 intend to find
a solution; if they success, \Metis replies a derivation in \TSTP
format of the proof, else the problem is counter feasible.

@@

\begin{align}
\begin{split}
\label{eq:unshunt-definition}
&\unshunt_{0}(φ) =\\
&\hspace{3mm}\begin{cases}
\unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃)
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\unshunt_{0}(φ₁ ⇒ φ₂) ∧ \unshunt_{0}(φ₁ ⇒ φ₃)
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃),\\
φ &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

@@

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
&\unshunt_{1}(φ, \zero) = φ\\
&\unshunt_{1}(φ, \suc(n)) =\\
&\hspace{3mm}\begin{cases}
\unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n)
  &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\unshunt_{1}(φ₁ ⇒ φ₂, n) ∧\ \unshunt_{1}(φ₁ ⇒ φ₃, n) &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{split}
\end{align}

@@

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
rank(φ₁) + c_{¬}  &\text{if } φ ≡ \neg φ₁, \\
\max{(rank(φ₁),rank(φ₂))} + c_{\square}
  &\text{if } φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

@@

\begin{equation}
\label{eq:unshunt-rank-definition}
\unshunt_{cm}(φ)=
\begin{cases}
\unshunt_{cm}(φ₃) + 2 & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\max(\unshunt_{cm}(φ₂),\unshunt_{cm}(φ₃)) + 1
  &\text{if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃),\\
0 &\text{otherwise.}
\end{cases}
\end{equation}

@@

\emph{Splitting a goal.}
The goal of a proof is processed by \Metis to generated new targets,
\Metis calls these new formulas \emph{subgoals}.
The conjunction of these subgoals is a proposition equivalent to the
original goal. Then, for each subgoal, in a \Metis' \TSTP derivation
we find a refutation.

Nevertheless, \Metis does not provide neither an explanation for the user about the process in its \TSTP derivations nor in its documentation.

In the following, we show how the $split$ function jointly with Theorem.~\ref{thm:thm-strip} prove the correctness of the process mentioned above.
First, we need to prove the correctness for the $\unshunt$ function in Eq~\ref{eq:unshunt-fixed}.

\begin{lemma}[\thmunshunt]
\label{lem:unshunt}
$Γ ⊢ \unshunt(φ)$, then $Γ ⊢ φ$.
\end{lemma}

@@

Now, let us consider the $split$ function that
yields the conjunction of subgoals from a proposition goal defined in Eq~\ref{eq:split-definition}.

\begin{equation}
\label{eq:split-definition}
% \mysplit(φ) =
\begin{cases}
\unshunt(\mysplit(φ₁)) ∧\ \unshunt(φ₁ ⇒ \mysplit(φ₂))  & \text{if }φ ≡ φ₁ ∧ φ₂,\\
      &\\[2.5mm]

\unshunt(¬ φ₁ ⇒ \mysplit(φ₂))       & \text{if }φ ≡ φ₁ ∨ φ₂,\\[2.5mm]

\unshunt(φ₁ ⇒ \mysplit(φ₂))         & \text{if }φ ≡ φ₁ ⇒ φ₂,\\[2.5mm]

\unshunt(φ₁ ⇒ \mysplit(φ₂))         & \text{if }φ ≡ φ₁ ⇔ φ₂,\\
∧\ \unshunt(φ₂ ⇒ \mysplit(φ₁))      &\\[2.5mm]

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))       & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂),\\[2.5mm]

\unshunt(\mysplit(¬ φ₁))            & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂),\\
∧\ \unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂))  &\\[2.5mm]

\unshunt(\mysplit(φ₁))              & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂),\\
∧\ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂))    &\\[2.5mm]

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))       & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂),\\
∧\ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))    &\\[2.5mm]

\unshunt(\mysplit(φ₁))              & \text{if }φ ≡ ¬ ¬ φ₁,\\
⊤                                   & \text{if }φ ≡ ¬ ⊥,\\
⊥                                   & \text{if }φ ≡ ¬ ⊤.
\end{cases}
\end{equation}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated Def.~\ref{eq:split-definition} to solve such a issue following the technique
described in Subsection~\ref{ssec:structural-recursion}.
Here and below, we refer to $split$ as the function after applying
the process mentioned above.

\begin{lemma}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\label{lem:lem-split}
\end{lemma}

@@

(φ ≡ φ₁ ∧ φ₂). We get a proof for each conjunct and using the introduction rule for conjunction, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ,n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁$.}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}$}
\UnaryInfC{Γ ⊢ φ₁}

\AxiomC{$Γ ⊢ \mysplit(φ,n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{$∧$-proj$₂$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{$⇒$-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ φ₂.$}
\end{bprooftree}
\end{equation*}

\vskip 3mm

(φ ≡ φ₁ ⇒ φ₂).
\begin{equation*}
\begin{bprooftree}
\AxiomC{}
\RightLabel{assume φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
\AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{weaken φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{⇒-elim}
\BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
\RightLabel{⇒-intro}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$.}
\end{bprooftree}
\end{equation*}
\vskip 3mm

(φ ≡ ¬ (φ₁ ∨ φ₂)). We show $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$. Using their conjunction by the introduction rule, and De Morgan Law, the results follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$,}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

\AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂.$}
\end{bprooftree}
\end{equation*}
\end{proof}

\begin{theorem}[thm-strip]
\label{thm:thm-strip}
 If $s_i$ are the subgoals for the goal φ, and
$\{ Γ ⊢ s_i \}_{i}$, then $Γ ⊢ φ$.
\end{theorem}

@@

\begin{align}
\begin{split}
conjunct(φ, ψ) =
\begin{cases}
  ψ &\text{if }φ ≡ ψ,\\
  ψ &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }conjunct(φ₁, ψ)≡ ψ,\\
  ψ &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }conjunct(φ₂, ψ)≡ ψ,\\
  φ &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

@@

\begin{proof}
When φ ≡ ψ, from $Γ ⊢ conjunct(φ, ψ)$ by normalization, we get $Γ ⊢ ψ$.
By using $subst$ theorem Eq.~\ref{eq:substitution-theorem}, we get the desire conclusion. If the proposition φ is a conjunction, and $ψ ≡ conjunct(φ_{i}, ψ)$
for some $i = 1,\ 2$, then,

\begin{equation}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ₁ ∧ φ₂$}
  \RightLabel{∧-proj$_{i}$}
  \UnaryInfC{$Γ ⊢ φ_{i}$}
  \UnaryInfC{$Γ ⊢ conjunct(φ_{i}, ψ)$}
  \AxiomC{$conjunct(φ_{i}, ψ) ≡ ψ$}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation}

Finally, in the latter case, by using the hypothesis, $Γ ⊢ φ$,
we complete the proof.
